---
title: "Markdown"
author: "Isabelle Smith"
date: "2025-01-21"
output: 
  html_document: 
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Test text: blah blah blah.

Two returns for new paragraph.

First line. (Two spaces...)  
Second line.

# Biggest Header

## Second Header




### (Multiple Returns...) Third Header

Use one set of asterisks for *italic text* and two sets for **bold text**. Use carets for ^superscript^ and tildes for ~subscript~. Two tildes makes ~~strikethrough~~.

- list item  
- (two spaces...) list item
  + indented list item


wow! a [link](www.google.com)!

Table | with | text
------|------|-----
lots  | of   | text
so    | many | rows


#### Fencing:

For `in-line` use a single tick mark, and for  

```
centered
```

use three.


#### In-Line Code:

Insert a calculation and only show output: `r 7 + pi`

#### Code Chunks:

```{r}

7 + pi

```


Adding `{r eval=TRUE, echo=FALSE}`:

```{r eval=TRUE, echo=FALSE}

7 + pi

```


#### LaTeX:

Single dollar signs: $a+b=c$

Double dollar signs: (new line)

$$a+b=c$$


##### Getting fancy:

sub- and superscripts...

$$H_A: \sigma^2 \neq 0$$

order of operations...

$$S=cA^z_1 + z_{2+x}$$


a classic...

$$f(x) = \frac{1}{\sqrt{2\pi \sigma}} e^{-\frac{(x-\mu)^2}{2 \sigma^2}}$$

summation...

$$\sum_{i=1}^n (y_i - \bar{y})^2$$

escaping characters...

$$\backslash$$

mixing...

$$P(\mbox{Species A} | weight>2.5) = a$$


```{r}

## comment

set.seed(777)

rpois(10,7)

```

