---
title: "Homework 10"
author: "Isabelle Smith"
date: "2025-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

## Libraries

```{r packages, message=FALSE, warning=FALSE}

##install.packages("tidytuesdayR")
library(tidytuesdayR)
library(tidyverse)
library(ggridges) # ridge plots
library(ggbeeswarm) # beeswarm plots
library(GGally) # parallel coordinates plots
library(ggpie) # pie charts
library(ggmosaic) # mosaic plots
library(scatterpie) # scatter pies on map
library(waffle) # for waffle plots
library(DescTools) # for converting table to long
library(treemap) # for tree maps

```

&nbsp;



## Work

Use your newly-developed `ggplot` chops to create some nice graphs from a dataset from the [TidyTuesday Dataset](https://github.com/rfordatascience/tidytuesday) project. You can make any combination of four plots from the list below:  

- Beeswarm Plot
- Ridgeline Plot
- Pie Charts
- Mosaic Plot
- Scatter Pie Plot
- Waffle Plot
- Tree Map
- 2-D Density Plot
- Dendrogram



&nbsp;

### 0. Data

I chose the [Palm Trees](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-03-18/readme.md) dataset.

```{r data, message=FALSE}

tuesdata <- tidytuesdayR::tt_load(2025, week = 11)
palmtrees0 <- tuesdata$palmtrees


palmtrees <- palmtrees0 |>
  dplyr::select(spec_name, acc_genus, palm_tribe, palm_subfamily,
         climbing, erect, max_stem_height_m, max_stem_dia_cm,
         average_fruit_length_cm, average_fruit_width_cm,
         fruit_size_categorical, fruit_shape, conspicuousness) |>
  drop_na() |>
  rename(fruit_conspic=conspicuousness)

```

```{r stats}

palmtrees |>
  dplyr::select(where(is.numeric)) |>
  summary()

palmtrees |>
  dplyr::select(!where(is.numeric)) |>
  mutate(across(everything(), function(x) length(unique(x)))) |>
  filter(row_number()==1) |>
  as.data.frame() |>
  `rownames<-`("n_unique") |>
  t()

```

```{r col}

d3_colors <- ggsci::pal_d3("category10")(10)

```



&nbsp;

### 1. Waffle Plot

```{r p1}

shape_table <- as.data.frame(table(shape=palmtrees$fruit_shape))

shape_colors <- c(d3_colors[1],
                  d3_colors[2],
                  d3_colors[10],
                  "darkolivegreen3", #d3_colors[9],
                  d3_colors[5],
                  d3_colors[3],
                  d3_colors[4])


ggplot(data=shape_table) +
  
  aes(fill = shape, values = Freq) +
  
  waffle::geom_waffle(n_rows = 25, size = 0.33, colour = "white") +
  
  scale_fill_manual(values=shape_colors) +
  
  coord_equal() +
  
  theme_void() +
  
  labs(title="Fruit Shapes of Palm Trees",
       fill=NULL) +
  
  theme(plot.title=element_text(hjust=0.5, vjust=4),
        plot.subtitle=element_text(hjust=0.5, vjust=4),
        plot.caption=element_text(vjust=-7),
        axis.title.x=element_text(vjust=-4),
        axis.title.y=element_text(vjust=5),
        plot.margin=margin(1, 1, 1, 1, "cm"),
        legend.position="top",
        legend.key.size = unit(0.02, 'npc')) +
  
  guides(fill = guide_legend(nrow = 1))

```



&nbsp;

### 2. Tree Map

```{r p2}

EC_table <- as.data.frame(table(Erect=palmtrees$erect,Climbing=palmtrees$climbing))


## increasing low counts for visualization purposes
EC_table_e <- EC_table |> mutate(Freq = case_when(Freq==0 ~ 0,
                                                 Freq < 10 ~ Freq + 20,
                                                 TRUE ~ Freq))

treemap(dtf=EC_table_e,
        index=c("Erect","Climbing"),
        vSize="Freq",
        type="index",
        title="Palm Tree Characteristics",
        fontsize.labels = c(11, 7),
        orce.print.labels= T,
        fontcolor.labels = c("black", "white"),
        align.labels = list(c("left", "top"), c("right", "bottom")),
        aspRatio=1.75)

```



&nbsp;

### 3. Scatterpie

```{r p3}

fruit_df_f <- palmtrees |>
  
  ## filtering size for visualization purposes
  filter(average_fruit_length_cm<10, average_fruit_width_cm<10) |>
  
  mutate(length_c = round(average_fruit_length_cm),
         width_c = round(average_fruit_width_cm),
         conspicuous = case_when(fruit_conspic=="conspicuous" ~ 1,
                                 fruit_conspic=="cryptic" ~ 0),
         cryptic = case_when(fruit_conspic=="cryptic" ~ 1,
                             fruit_conspic=="conspicuous" ~ 0)) |>
  
  group_by(length_c,width_c) |>
  
  summarize(n = n(),
            conspicuous=sum(conspicuous),
            cryptic=sum(cryptic),
            .groups="drop")



ggplot(data=fruit_df_f) + 
  scatterpie::geom_scatterpie(
    aes(x=width_c, y=length_c), 
    pie_scale=2,
    cols=c("conspicuous", "cryptic")) + 
   coord_equal() +
  scale_fill_manual(values=c("coral","turquoise")) +
  labs(title="Distribution of Palm Tree Fruit",
       x="Average Width",
       y="Average Length",
       fill="Color Type") +
  
  theme(plot.title=element_text(hjust=0.5, vjust=4),
        plot.subtitle=element_text(hjust=0.5, vjust=4),
        plot.caption=element_text(vjust=-7),
        axis.title.x=element_text(vjust=-4),
        axis.title.y=element_text(vjust=5),
        plot.margin=margin(1, 1, 1, 1, "cm"),
        legend.position="right")

```



&nbsp;

### 4.

```{r p4}

size_df <- palmtrees |>
  
  mutate(average_fruit_size = average_fruit_length_cm * average_fruit_width_cm,
         
         ## reversing order to be alphabetical in visualization
         palm_subfamily_r = fct_rev(palm_subfamily))


size_df_f <- size_df |>
  
  ## filtering size for visualization purposes
  filter(average_fruit_size < 35)



ggplot(data=size_df_f) +
  
  aes(x=average_fruit_size,y=palm_subfamily_r,fill=palm_subfamily_r) + 
  
  ggridges::geom_density_ridges() +
  
  ggridges::theme_ridges() +
  
  ggsci::scale_fill_observable() +
  
  labs(title="Distribution of Palm Tree Fruit by Subfamily",
       x="Average Fruit Size",
       y="Subfamily") +
  
  theme(plot.title=element_text(hjust=0.5, vjust=4),
        plot.subtitle=element_text(hjust=0.5, vjust=4),
        plot.caption=element_text(vjust=-7),
        axis.title.x=element_text(hjust=0.5, vjust=-4),
        axis.title.y=element_text(hjust=0.5, vjust=5),
        plot.margin=margin(1, 1, 1, 1, "cm"),
        legend.position="none")

```


