---
title: "Homework 10"
author: "Isabelle Smith"
date: "2025-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

&nbsp;

## Libraries

```{r, message=FALSE}

library(tidyverse)
##install.packages("tidytuesdayR")

```

&nbsp;



## Work

Use your newly-developed `ggplot` chops to create some nice graphs from a dataset from the (TidyTuesday Dataset)[https://github.com/rfordatascience/tidytuesday] project. You can make any combination of four plots from the list below:  

- Beeswarm Plot
- Ridgeline Plot
- Pie Charts
- Mosaic Plot
- Scatter Pie Plot
- Waffle Plot
- Tree Map
- 2-D Density Plot
- Dendrogram



&nbsp;

### Data

I chose the (Palm Trees)[https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-03-18/readme.md] dataset.

```{r, message=FALSE}

tuesdata <- tidytuesdayR::tt_load(2025, week = 11)
palmtrees0 <- tuesdata$palmtrees


palmtrees <- palmtrees0 |>
  select(spec_name, acc_genus, palm_tribe, palm_subfamily,
         climbing, erect, max_stem_height_m, max_stem_dia_cm,
         average_fruit_length_cm, average_fruit_width_cm,
         fruit_size_categorical, fruit_shape, conspicuousness) |>
  drop_na() |>
  rename(fruit_conspic=conspicuousness)

```

```{r}

palmtrees |>
  select(where(is.numeric)) |>
  summary()

palmtrees |>
  select(!where(is.numeric)) |>
  mutate(across(everything(), function(x) length(unique(x)))) |>
  filter(row_number()==1) |>
  as.data.frame() |>
  `rownames<-`("n_unique") |>
  t()

```



&nbsp;

### 1. Waffle Plot

```{r}

shape_table <- as.data.frame(table(shape=palmtrees$fruit_shape))

d3_colors <- ggsci::pal_d3("category10")(10)

shape_colors <- c(d3_colors[1],
                  d3_colors[2],
                  d3_colors[9],
                  d3_colors[10],
                  d3_colors[5],
                  d3_colors[3],
                  d3_colors[4])


ggplot(data=shape_table) +
  
  aes(fill = shape, values = Freq) +
  
  waffle::geom_waffle(n_rows = 25, size = 0.33, colour = "white") +
  
  scale_fill_manual(values=shape_colors) +
  
  coord_equal() +
  
  theme_void() +
  
  labs(title="Shapes of Palm Tree Fruit",
       fill=NULL) +
  
  theme(plot.title=element_text(hjust=0.5, vjust=4),
        plot.subtitle=element_text(hjust=0.5, vjust=4),
        plot.caption=element_text(vjust=-7),
        axis.title.x=element_text(vjust=-4),
        axis.title.y=element_text(vjust=5),
        plot.margin=margin(1, 1, 1, 1, "cm"),
        legend.position="top",
        legend.key.size = unit(0.02, 'npc')) +
  
  guides(fill = guide_legend(nrow = 1))

```



&nbsp;

### 2.

```{r}

```



&nbsp;

### 3.

```{r}

```



&nbsp;

### 4.

```{r}

```


